Reduced Nucleotide Diversity on a Plant Y Chromosome Following Recent Recombination Suppression

### 2016-08-01 manuscript.tex
fixed bibliography issue
minor edits

%\usepackage[colorinlistoftodos]{todonotes} % comments in margins
%\setlength{\marginparwidth}{1.25cm}
\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
\newcommand{\X}{\textcolor{red}{\bf X\,}}

@@

Of particular relevance to the present study, \textit{R. hastatulus} populations throughout their geographical range (southern USA) have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}. This is important because female-biased sex ratios are predicted to affect levels of sex chromosome diversity \citep{ellegren2009}. Here, we focus on the $Y_{1}$ chromosome that is shared between the chromosomal races of \textit{R. hastatulus}, and investigated patterns of nucleotide diversity on X, Y, and autosomal chromosomes. Our goal was to determine whether the levels of neutral diversity on these chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection (background selection and/or selective sweeps) has likely had a greater impact on polymorphism levels. To do this, we used empirically-estimated population genetic parameters to conduct forward-time simulations of purifying and positive selection and evaluated the relative support for these models in explaining the levels of diversity observed in our data.

@@

\subsection*{Simulations of Background Selection and Selective Sweeps}
To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations..a

\josh{
\textit{to do}
\begin{itemize}
\item \textit{describe parameters used and justify}
\item \textit{do both purifying and positive selection sims}
\end{itemize}
}

We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}. To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data. To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$. We then calculated the average $\theta$ over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure). Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).

>>>>

%\usepackage[colorinlistoftodos]{todonotes} % comments in margins
%\setlength{\marginparwidth}{1.25cm}
%\definecolor{flame}{rgb}{0.89, 0.35, 0.13}
%\newcommand{\josh}[1]{\textcolor{flame}{\emph{\scriptsize {#1}}}}
%\newcommand{\citex}{\textcolor{red}{\bf (CITE)\,}}
%\newcommand{\out}[1]{\textcolor{red}{#1}} %outline placeholder text
\newcommand{\X}{\textcolor{red}{\bf X\,}}

@@

Of particular relevance to the present study, \textit{R. hastatulus} populations throughout their geographical range (southern USA) have also been found to consistently exhibit female-biased sex ratios, with a mean sex ratio of $N_{m}/N_{f}=0.6$ \citep{pickup2013influence}. This is important because female-biased sex ratios are predicted to affect genetic diversity on sex chromosomes and autosomes \citep{ellegren2009}. Here, we investigated the factors affecting nucleotide diversity on X, Y, and autosomal chromosomes. Our goal was to determine whether levels of diversity on these chromosomes were jointly consistent with predictions from models of sex ratio bias and variance in reproductive success, or whether linked selection has been the dominant factor affecting patterns of neutral polymorphism on sex chromosomes.

@@

\subsection*{Simulations of Background Selection and Selective Sweeps}
To investigate whether our observed level of Y-chromosome diversity could be explained by background selection, we conducted forward time simulations.....We modeled both purifying and positive selection using the forward simulation software SFSCODE \citep{hernandez2008flexible}. To make the simulation output more comparable to our data, we used estimates of $\theta$ from our polymorphism data as a simulation input parameter, and sampled 8 Y chromosomes per simulation from 24 populations to match out empirical data. To consider the effects of selection against deleterious alleles, we sampled selective coefficients from a gamma distribution with shape parameter 0.1, mean selection coefficient $s=-0.01$, neutral mutation rate $\mu = 0.005$, deleterious mutation rate $U=0.01$ (diploid), and dominance term $h=1$. We then calculated the average $\theta$ over all simulations, and obtained the Y/A ratio of neutral diversity under the model and compared this to our empirical data (\X new figure). Confidence intervals were calculated from bootstrapping 10000 replicate simulations  using the BCa method (Efron 1987) implemented in the Boot package in R (Canty and Ripley 2012; R Core Team 2011).
